---
title: "Assignment 5"
author: "Ian Tongs"
date: "27/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
library(forecast)
```

# Section 8.8 Buildup Exercises:

## Question 2:

```{r echo = TRUE, message=FALSE, warning=FALSE}
simple_exponential_smoothing <- function(y, alpha, level){
  t <- length(y)
  l <- numeric(t + 1) # numeric creates a vector of zeroes of length t+1 (+1 for the forecast)
  l[1] <- level
  for (i in 2:(t+1)) {
    l[i] <- alpha * y[i-1] + (1 - alpha) * l[i - 1]
  }
  fcast <- l[t + 1]
  return(fcast)
}

y <- c(11,13,12,10,15,16,14)

testfit <- ses(y, h=1)

alpha <- testfit$model$par[1]
level_0 <- testfit$model$par[2]

simple_exponential_smoothing(y, alpha, level_0)

testfit

```




## Question 3:

```{r echo = TRUE, message=FALSE, warning=FALSE}
SSE_SES <- function(y, par){
  alpha <- par[1]
  level <- par[2]
  t <- length(y)
  l <- numeric(t + 1) # numeric creates a vector of zeroes of length t+1 (+1 for the forecast)
  l[1] <- level
  for (i in 2:(t+1)) {
    l[i] <- alpha * y[i-1] + (1 - alpha) * l[i - 1]
  }
  SSE <- 0
  for (i in 1:t) {
    SSE <- SSE + (y[i] - l[i])^2
  }
  return (SSE)
}

y <- c(11,13,12,10,11)
#alpha <- 0.25
#level <- 10

#SSE_SES(y, alpha, level)

optimised <- optim(par=c(0.5,y[1]), fn=SSE_SES, y=y)
optimised$par

testfit <- ses(y, h=1)
testfit$model$par



SSE_SES_aux2 <- function(y, par){
  error <- 0
  SSE <- 0
  alpha <- par[1]
  l0 <- par[2]
  y_hat <- l0
 
    for(index in 1:length(y)){
  error <- y[index] - y_hat
    SSE <- SSE + error^2
  
    y_hat <- alpha*y[index] + (1 - alpha)*y_hat 
  }
 
  return(SSE)
}

optimised <- optim(par=c(0.5,y[1]), fn=SSE_SES_aux2, y=y)
optimised$par

# ts_y <- as_tsibble(y)
# 
# fit <- ts_y %>%
#   model(ETS(y ~ error("A") + trend("N") + season("N"), opt_crit = "mse"))
# fc <- fit %>%
#   forecast(h = 5)

```



    t <- length(y)
    l <- numeric(t + 1) # numeric creates a vector of zeroes of length t+1 (+1 for the forecast)
    l[1] <- level
    for (i in 2:(t+1)) {
      l[i] <- alpha * y[i-1] + (1 - alpha) * l[i - 1]
    }
    fcast <- l[t + 1]
    return(fcast)



    alpha <- par[1]
    level <- par[2]
    t <- length(y)
    l <- numeric(t + 1) # numeric creates a vector of zeroes of length t+1 (+1 for the forecast)
    l[1] <- level
    for (i in 2:(t+1)) {
      l[i] <- alpha * y[i-1] + (1 - alpha) * l[i - 1]
    }
    SSE <- 0
    for (i in 1:t) {
      SSE <- SSE + (y[i] - l[i])^2
    }
    return (SSE)

# Assignment Proper:


```{r echo = TRUE, message=FALSE, warning=FALSE}
simple_exponential_smoothing_function <- function(y) {
  
  # Auxiliary function for smoothing with a given alpha an initial level
  simple_exponential_smoothing_auxiliary <- function(y, alpha, level){
    t <- length(y)
    l <- numeric(t + 1) # numeric creates a vector of zeroes of length t+1 (+1 for the forecast)
    l[1] <- level
    for (i in 2:(t+1)) {
      l[i] <- alpha * y[i-1] + (1 - alpha) * l[i - 1]
    }
    fcast <- l[t + 1]
    return(fcast)
  }
  
  # Returns the value of the SSE for the SES with par = c(alpha, level) )
  SSE_SES_aux <- function(y, par){
    alpha <- par[1]
    level <- par[2]
    t <- length(y)
    l <- numeric(t + 1) # numeric creates a vector of zeroes of length t+1 (+1 for the forecast)
    l[1] <- level
    for (i in 2:(t+1)) {
      l[i] <- alpha * y[i-1] + (1 - alpha) * l[i - 1]
    }
    SSE <- 0
    for (i in 1:t) {
      if (alpha>0) {
        SSE <- SSE + (y[i] - l[i])^2
      } else {
        SSE <- SSE + (y[i])^2
      }
    }
    return (SSE)
  }
  
  # Optimise the parameters for the forecast:
  optimised <- optim(par=c(0.6, y[1]), fn=SSE_SES_aux, y=y)
  alpha <- optimised$par[1]
  level_0 <- optimised$par[2]
  
  print(optimised$par)
  print(SSE_SES_aux(y, optimised$par))
  print(SSE_SES_aux(y, fit1$model$par))
  
  # Get the one step ahead forecast
  pred_val <- simple_exponential_smoothing_auxiliary(y, alpha, level_0)
  
  # As this is an SES model, all ten predictions are equal, so:
  forecast_vector <- numeric(10)
  for (i in 1:10) {
    forecast_vector[i] <- pred_val
  }
  
  # Return the forecast vector
  return (forecast_vector)
}
  
y <- c(11,13,12,10,15,16,14)
  

simple_exponential_smoothing_function(y)       # Choose a y to get this to work


# # create a tsibble w/o a key
# test_y <- tsibble(
#   date = as.Date("2017-01-01") + 0:6,
#   value = c(11,13,12,10,15,16,14)
# )
# 
# # Estimate parameters
# fit <- test_y %>%
#   model(ETS(value ~ error("A") + trend("N") + season("N"), opt_crit = "mse"))
# fc <- fit %>%
#   forecast(h = 10)
# 
# report(fit)
# 
# fc

fit1 <- ses(y, h=1)
fit1
fit1$model$SSE
fit1$model$par

```



# To Do:
- Q2 AND Q3 from 8.8 are good for the assignment
- want: want a time series y, an alpha and a level and then does the exponential smothing
- make a script version